---
title: "Functions"
author: "Erin McLean"
date: "2/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(roxygen2)
library(mytools)
```
#Setting up some dummy data

```{r}
airtemps<- c(212, 30.3, 78, 32, 0) 
celsius1<-(airtemps[1]-32)*(5/9)
celsius1
celsius2<-(airtemps[2]-32)*(5/9)
celsius2
celsius3<-(airtemps[3]-32)*(5/9)
celsius4<-(airtemps[4]-32)*(5/9)
celsius5<-(airtemps[5]-32)*(5/9)
```

#Defining temperature conversion function

"#' " is meant to be the start of documentation for a function that's meant to be saved. Start with what the function does, starting with a verb. Should be short, because it'll be the title.

Then add a line and add more documentation.

Telling it the names of the parameters that it needs for input. First, F to C:

```{r}

#' convert degrees F to C
#' 
#' we use the simple formala for temp conversion to convert.
#' one can find this in any intro to physics book or on google.
#' @param fahr the temperature to be converted in degrees F
#' @return the temperature converted to C
#' @examples 
#' fahr_to_cel(32)
#' fahr_to_cel(c(212,100,32))

fahr_to_cel<-function(fahr) {
  #stopifnot(is.numeric(fahr))
  cel<-(fahr-32)*(5/9)
  return(cel)
}
```

If you don't think you're going to use the functionality again, put your code in a loop to avoid repeating yourself.

```{r}
cel4<-fahr_to_cel(airtemps[4])
cel4

celsius<-fahr_to_cel(airtemps)
celsius
```
calling it that way only works because it's in the same directory. building a package makes it a lot easier to source things that doesn't depend on your particular computer setup of files and folders.

```{r}
source('Functions.R')

cel_to_fahr<-function(cel) {
  fahr<-(cel*(9/5))+32
  return(fahr)
}
```
fahrenheit<-cel_to_fahr(celsius)
fahrenheit


```{r}
environment_info("hey NCEAS")
```

debug(fahr_to_cel)
install.packages("blogdown")

