---
title: "Data Cleaning Exercise"
author: "Erin McLean"
date: "2/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading in Packages

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
```

#Loading in Data

```{r}
daily_orig<-read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", stringsAsFactors = FALSE)
head(daily_orig)
```

#Filtering to just Sockeye & Bristol Bay

```{r}
BBsockeye_catch<-daily_orig %>% 
  filter(Species=="Sockeye",SASAP.Region=="Bristol Bay") %>% 
  select(-Method,-Latitude,-Longitude,-Source)
head(BBsockeye_catch)
```

#Separating the Date column into 3, grouping by year and location, getting the sum at each location and each year 

```{r}
byyear_BBsockeye <-  
  separate(BBsockeye_catch,sampleDate,
           into=c("Year","Month","Day"),
           sep="-",
           remove=T) %>% 
  group_by(Year,Location) %>% 
  summarise(Escapement=sum(DailyCount)) 
print(byyear_BBsockeye)
```

#Loading in 2nd set of Data

```{r}
goals_orig<-read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/knb.92014.1", stringsAsFactors = FALSE)
head(goals_orig)
```

#Filtering to just Sockeye & Bristol Bay, changing lower limit to integer

```{r}
BBsockeye_goal<-goals_orig %>% 
  filter(Species=="Sockeye",Region=="Bristol Bay") %>% 
  select(System, Lower,Upper,Initial.Year) %>% 
  mutate(Lower=as.integer(Lower))

head(BBsockeye_goal)
```

#Data QA

```{r}
i<-which(is.na(BBsockeye_goal$Lower))
BBsockeye_goal[i,]
```

#Joining the tables

```{r}
goals_joined<-left_join(byyear_BBsockeye,BBsockeye_goal,
                        by=c("Location"="System")) %>%
  filter(Year>=Initial.Year) %>% 
  filter(Escapement>=Lower)
print(goals_joined)
```

#Are these data meeting the lower goals?

```{r}
goals_lower<-filter(goals_joined, Escapement>=Lower) %>% 
  select(-Lower,-Initial.Year) 

print(goals_lower)
```


#Are these data meeting the upper goals?

```{r}
goals_upper<-filter(goals_joined, Escapement<=Upper) %>% 
  select(-Lower,-Initial.Year) 

print(goals_upper)
```


#Are these data meeting both goals?

```{r}
goals<-filter(goals_joined, Lower<=Escapement & Escapement>=Upper) %>% 
  select(-Initial.Year) 

print(goals)
```