---
title: "Spatial_Data"
author: "Erin McLean"
date: "2/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(scales)
library(leaflet)
library(ggmap)
```

Plot in base R helps you get a quick view of your data. Do you need to change your projection?
st_crs tells you what coordinate reference system that you're in
EPSG codes: 
4326 (WGS84)
3338 (Alaska Albers)
3857 (Pseudo Mercator)
```{r}
ak_regions<-read_sf("shapefile_demo_data/ak_regions_simp.shp")
class(ak_regions)
plot(ak_regions)
st_crs(ak_regions)
```

often you need to have different versions of your data in different projections

sf works well with the tidyverse but your geometry column is sticky - it stays around because that's where your spatial data lives.
```{r}
ak_regions_3338<-ak_regions %>% 
  st_transform(crs=3338)

nrow(ak_regions_3338)
summary(ak_regions_3338)

plot(ak_regions_3338)
```

Reading in population data

```{r}
pop<-read.csv("shapefile_demo_data/alaska_population.csv",stringsAsFactors = F)
head(pop)
```
Can do a spatial join but both must be of class sf.
Remove determines whether to keep the lat and lang - in this case, you want to keep them.

```{r}
pop_4326<-st_as_sf(pop,
                   coords=c('lng','lat'),
                   crs=4326,
                   remove=F)

pop_3338<-pop_4326 %>% 
  st_transform(crs=3338)
```

Do the spatial join of the two datasets
St within asks the question which of the points belongs inside which of the polygons
If we had lines or other polygons, we'd be looking at ST intersect - which line intersects which polygon.
The default is the left join.

```{r}
pop_joined<-st_join(pop_3338,ak_regions_3338, join=st_within)
plot(pop_joined)
```

Now we want to calculate the pop by region.

Now this is a vector of cities in each region...as opposed to a polygon. Total pop info should be associated with the region rather than the points.

```{r}
pop_region<-pop_joined %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarise(total_pop=sum(population))

#now we can make the geometry column unsticky - we don't need the point data, we want it to be associated with a polygon. Drop the geometry first. That's what the as.data.frame does

head(pop_region)
```
Create new sf that has pop and region info

Can do regular left joins on spatial data as long as the spatial data is the first arg and the second join is the regular data frame
```{r}
pop_region_3338<-left_join(ak_regions_3338,pop_region)

plot(pop_region_3338)
```
this unions the geography

if you don't want to do that, add an arg to summarize function
```{r}
pop_mgmt_3338<-pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarise(total_pop=sum(total_pop),do_union=F)

plot(pop_mgmt_3338["total_pop"])
```

write files to disk, good when it takes a long time to process

```{r}
write_sf(pop_region_3338, "shapefile_demo_data/ak_region_population.shp",
         delete_layer = T)
```

Make some maps!

This makes geospatial analysis reproducible. Don't add axes for geospatial data. Instead you're going ot layer different vars on to your map.

```{r}
ggplot(pop_region_3338)+
  geom_sf(mapping=aes(fill=total_pop))+
  theme_bw()+
  labs(fill="Total Population")+
  scale_fill_continuous(low="turquoise1",
                        high="midnightblue",
                        labels=comma)
```

```{r}
rivers_3338 <- read_sf("shapefile_demo_data/ak_rivers_simp.shp")
st_crs(rivers_3338)
```

This is how you add layers. Drawn in the order in which you give them.
```{r}
ggplot() +
  geom_sf(data = pop_region_3338, aes(fill = total_pop)) +
  geom_sf(data=rivers_3338,aes(size=StrOrder),color="black")+
  geom_sf(data=pop_3338,aes(), size=.5)+
  scale_size(range=c(.01,.2),guide=F)+
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high =  "firebrick", labels = comma)
```

3857 is the projection system used by google maps, so if you want to use those, you have to transform your datasets

```{r}
pop_3857<-pop_3338 %>% 
  st_transform(crs=3857)
```

CRS from the tile server doesn't match the bounding box. This function fixes it.
```{r}
# Define a function to fix the bbox to be in EPSG:3857
# See https://github.com/dkahle/ggmap/issues/160#issuecomment-397055208
ggmap_bbox_to_3857 <- function(map) {
  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
  # and set the names to what sf::st_bbox expects:
  map_bbox <- setNames(unlist(attr(map, "bb")), 
                       c("ymin", "xmin", "ymax", "xmax"))
  
  # Coonvert the bbox to an sf polygon, transform it to 3857, 
  # and convert back to a bbox (convoluted, but it works)
  bbox_3857 <- st_bbox(st_transform(st_as_sfc(st_bbox(map_bbox, crs = 4326)), 3857))
  
  # Overwrite the bbox of the ggmap object with the transformed coordinates 
  attr(map, "bb")$ll.lat <- bbox_3857["ymin"]
  attr(map, "bb")$ll.lon <- bbox_3857["xmin"]
  attr(map, "bb")$ur.lat <- bbox_3857["ymax"]
  attr(map, "bb")$ur.lon <- bbox_3857["xmax"]
  map
}
```

```{r}
bbox <- c(-170, 52, -130, 64)   # This is roughly southern Alaska
ak_map <- get_stamenmap(bbox, zoom = 4)
ak_map_3857 <- ggmap_bbox_to_3857(ak_map)
```

```{r}
ggmap(ak_map_3857) + 
  geom_sf(data = pop_3857, aes(color = population), inherit.aes = F) +
  scale_color_continuous(low = "khaki", high =  "firebrick", labels = comma)
```

